### 2013-08-15 cover.tex

\thispagestyle{empty}

\hfill ORNL/TM-2013/151

\vspace{3em}

\begin{center}
National Center for Computatational Sciences
\end{center}

\vspace{3em}

\begin{center}
\textbf\textsc{\Large Ceph Parallel File System Evaluation Report}\\
\rule{5in}{1pt}%
\end{center}

\vspace{0.5in}

\begin{center}
\begin{tabular}{ c c }
Feiyi Wang  &  Mark Nelson \\ 
ORNL/NCCS & Inktank Inc. \\
\end{tabular}
\end{center}

\vspace{0.25in}

\begin{center}
Other contributors: \\
\vspace{1em}
Sarp Oral\\
Doug Fuller\\
Scotty Atchley\\
Blake Caldwell\\
James Simmons\\
Brad Settlemyer\\
Jason Hill\\
Sage Weil (Inktank Inc.)
\end{center}

\vfill

\begin{center}
Prepared by \\
OAK RIDGE NATIONAL LABORATORY\\
Oak Ridge, Tennessee 37831-6283 \\ 
managed by \\
UT-BATTELLE, LLC \\
for the \\
U.S. DEPARTMENT OF ENERGY\\ 
under contract DE-AC05-00OR22725
\end{center}

\pagebreak