### 2013-08-21 baseline.tex
Did a quick breakup from original report, resize figures to fit to IEEE format.

\section{Baseline Performance}
\label{sec:block-io}

\subsection{Block I/O over Native IB}

We first established baseline performance by measuring block I/O performance.
At the block-level, with each LUN configured as a RAID6 8+2 array, we had the
following results as shown in Table~\ref{tbl:block-io-baseline}.

\begin{table}[htb]
\centering
\caption{Block I/O performances on single LUN and single host}
\label{tbl:block-io-baseline}

\begin{tabular}{| l | l |}
    \hline
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\
    Single host with four SAS LUNs & ~ 2.8 GB/s \\
    Single host with seven SATA LUNs & ~ 2.6 GB/s \\
    \hline
\end{tabular}
\end{table}

Single host in this table refers to one of four tick-oss nodes. Four tick-oss
nodes drive the SFA10K backend storage. Overall, we observe that the entire
system can perform at 11 GB/s, compared to DDN SFA10K's theoretical maximum of
12 GB/s.

\subsection{IP over IB}

Ceph uses the BSD Sockets interface in \texttt{SOCK\_STREAM} mode (i.e., TCP).
Because our entire testbed is IB-switched, we used IP over IB (IPoIB) for
networking\footnote{As of writing of this report, Inktank is investigating using
rsockets to improve performance with IB fabric}. Through simple
\verb!netperf! tests, we observed that a pair of hosts connected by IB QDR using
IPoIB can transfer data at 2.7 GB/s (the hosts are tuned per recommendations
from Mellanox). With all four hosts (OSD servers), we expect the aggregate
throughput to be in the neighborhood of 10 GB/s.

Unfortunately there was not enough time to do more detailed analysis of the
network performance. However, as we observed later, RADOS is performing no
more than 8 GB/s driven by four server hosts. This confirms that we have
provisioned enough network bandwidth. In other words, IP over IB is not a
bottleneck in this case.