### 2013-08-28 baseline.tex
changes to the basline section

\section{Establishing A Baseline Performance}
\label{sec:baseline}

\subsection{Block I/O over Native IB}

The lowest layer in the system are the exposed LUNs to the server hosts from
the RAID controllers.  These LUNs are mostly configured as a RAID6 8+2 array (8
data disks and 2 \textit{P} and \textit{Q} disks). In our earlier tests we
observed that a 7.2K RPM SAS disk can perform at ~140 MB/s for 128 kB
sequential I/O requests and a 7.2K RPM SATA disk can do ~36 MB/s. For these
tests, all disk caches were turned off. Therefore, in 8 data disk RAID group
with 1 MB sequential I/O requests (each disks sees 128 kB chunks of the
request) the expected aggregate performance would be ~1.12 GB/s for a SAS RAID
group and ~288 MB/s for a SATA RAID group, if there are no caches along the I/O
path.  With the caches turned on, especially the write-back cache on the DDN
RAID controller, we observe a significant performance improvement. In our
day-to-day HPC operations we run our storage systems with write-back cache
turned on on RAID controllers, if and only if, there is a method of
cache-mirroring between the two active-active RAID controllers to recover from
a single controller failure scenario. With the write-back cache on we observe
~1.4 GB/s per RAID 6 8+2 SAS group and ~955 MB/s per RAID 6 8+2 SATA group, for
1 MB sequential I/O. These two performance numbers will be used as a baseline
in our study and they are presented in Table~\ref{tbl:block-io-baseline}.  For
these tests, the 8+2 RAID 6 groups are exported to the server hosts using SCSI
RDMA Protocol (SRP) over IB protocol (which is the default for our normal HPC
configurations). For a SATA disk group this translates into roughly 120
MB/s/disk and for a SAS disk group it is 175 MB/s. As can be seen, SATA RAID
groups benefit from caching more than SAS groups.

For above tests the RAID groups are exported over physical 4x IB QDR links.
These links can provide up to 40 Gbits/s signalling rate and 32 Gbits/s actual
data rate (reduction is due to 8/10 encoding scheme). Therefore, each IB QDR
link can perform at 3.2 GB/s, while in practice we see around 3 GB/s as the
best case.

As mentioned earlier, we had 280 SATA disks and 200 SAS disks. Organizing these
into RAID 6 8+2 groups will yield 28 SATA and 20 SAS RAID 6 8+2 groups. We had
4 server hosts in our testbed and evenly distributing these on to 4 server
hosts translates into 8 SATA LUNs and 4 SAS LUNs per server host (when a RAID
group is exported to a server host, it is labeled as a LUN from the server host
point of view).

Next, we established the baseline performance for a single server host. We
exercised all seven SATA LUNs concurrently form a single server host at the
block-level. This resulted in an aggregate performance of roughly 2.6 GB/s.
Repeating the same test using only four SAS LUNs yielded around 2.8 GB/s
aggregate performance for 1 MB sequential I/O. These two numbers will be used
as single server host aggregate baseline performance for this study. These are
also presented in presented in Table~\ref{tbl:block-io-baseline}.

It is worth mentioning that we used our in-house developed synthetic benchmark,
\textit{fair-lio} for all our block-level testing.  fair-lio has better
sequential I/O characteristics compared to commonly used XDD benchmark and it
is asynchronous and based on the Linux libaio library.

\begin{table}[htb]
\centering
\caption{Block I/O performances on single LUN and single host}
\label{tbl:block-io-baseline}

\begin{tabular}{| l | l |}
    \hline
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\
    Single host with four SAS LUNs & ~ 2.8 GB/s \\
    Single host with seven SATA LUNs & ~ 2.6 GB/s \\
    \hline
\end{tabular}
\end{table}

When exercising all 28 SATA LUNs from all four server hosts in parallel, we
observed an 11 GB/s aggregate performance. Same level of performance was
observed using all 20 SAS LUNs from four server hosts in parallel. Comparing
these to the advertised peak performance of 12 GB/s of the DDN SFA10K, we
concluded that our test setup is well configured to drive the backend disk
system at full performance and we are limited by the RAID controller
performance. Going forward in this study, 11 GB/s will be used the peak
baseline for the entire test configuration.

\subsection{Establishing an IP over IB Baseline Performance}

Ceph uses the BSD Sockets interface in \texttt{SOCK\_STREAM} mode (i.e., TCP).
Because our entire testbed is built on IB QDR fabric (including client to
server host connections over a 108-port Mellanox IB QDR switch), we used IP
over IB (IPoIB) for networking\footnote{As of writing of this report, Inktank
is investigating using rsockets to improve performance with IB fabric}. Through
simple \verb!netperf! tests, we observed that a pair of hosts connected by IB
QDR using IPoIB can transfer data at 2.7 GB/s (the hosts are tuned per
recommendations from Mellanox). With all four server hosts (OSD servers), we
expect the aggregate throughput to be in the neighborhood of 11 GB/s.

Unfortunately there was not enough time to do more detailed analysis of the
network performance. However, as we observed later, RADOS is performing no
more than 8 GB/s driven by four server hosts. This confirms that we have
provisioned enough network bandwidth. In other words, IP over IB is not a
bottleneck in this case.