### 2013-08-29 abstract.tex
fix the flow and minor typos

\begin{abstract}

%Ceph is designed to be a reliable and scalable fault tolerant parallel file
%system with a number of intriguing features.

Ceph is an open-source and emerging parallel distributed file and storage
system technology. It is designed to provide file system-level and object-level
access to data. By design, Ceph assumes running on unreliable and commodity
storage and network hardware and provides reliability and fault-tolerance
through pre-calculated and controlled object placement and data replication.

In collaboration with its major developer, Inktank Inc, we evaluated the Ceph
technology for scientific high-performance computing (HPC) environments.  This paper
presents our evaluation methodology, experiments, results and observations from
mostly parallel I/O performance and scalability perspectives.  Our work made
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment like ours using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should also benefit both Ceph and HPC communities at
large. Throughout the evaluation, we observed that Ceph still is an evolving
technology under fast-paced development and showing great promises.

\end{abstract}