### 2013-09-20 conclusion.tex
Brad's attempt to cut to 5 pages
So here is most of my attempt to cut this paper to 5 pages.
I amde it to about 5.7 pages.  I could go after it some more,
but this is a good first draft point to check and see if we need
to add back anything I took out.

I also added the IEEEabbrv.bib file since I couldn't build
without it.

\section{Observations and Conclusions}
\label{sec:conclusion}

%Ceph is built on the assumption that the underlying hardware components are
%unreliable, with little or no redundancy and failure detection capability.
%This assumption is not valid for high-end HPC centers like OLCF. We have
%disabled replication for pools, haven't measured and quantified processing
%overhead, and we do not know yet if this would be significant. Investigating
%this remains as a future work. 
In our early tests, we experienced large performance swings during different
runs, low read performance when transfer size is small, and I/O errors tend to
happen when system is under stress (more clients and large transfer sizes).
However, with systematic performance engineering and development efforts, we
have seen a steady improvement through different releases. As of now, the Ceph
system on our testbed is able to perform close to 80\% of raw hardware
capability at RADOS level and close to 70\% at file system level (after
accounting for journaling overheads) . 
%This is still no comparison to Lustre
%yet, but by no means a small feat for such a \textit{young} technology. It is,
%in fact, a very respectable level of performance.

Ceph is built on the assumption that the underlying hardware components are
unreliable, with little or no redundancy and failure detection capability.
This assumption is not valid for high-end HPC centers like OLCF.
Ceph performs \textbf{metadata + data} journaling, which is appropriate for host
systems that have locally attached disks rather than DDN SFA10K-like hardware,
where the backend LUNs are exposed as block devices through IB over SRP
protocol. The journaling write requires twice the bandwidth compared to
Lustre-like meta data-only journaling mechanism. For Ceph to be viable in
large-scale capability HPC environments like OLCF, journaling operations need
to support HPC storage hardware.

%The current design on journaling write does present a challenge in
%our IB-switched storage hardware. As BTRFS and other backend file system
%mature, we are seeing promising signs for Ceph to take advantage for a
%better journaling design.