### 2013-10-02 abstract.tex
cutting

\begin{abstract}

%Ceph is designed to be a reliable and scalable fault tolerant parallel file
%system with a number of intriguing features.

Ceph is an emerging open-source parallel distributed file and storage
system. By design, Ceph is designed to leverage unreliable and commodity
storage and network hardware, and provides reliability and fault-tolerance
via controlled object placement and data replication.  
%Here we evaluate Ceph
%for scientific high-performance computing (HPC) environments.  
This paper
presents our file and block I/O performance and scalability evaluation of Ceph for
scientific high-performance computing (HPC) environments.
% methodology, experiments, results and observations from
%mostly parallel I/O performance and scalability perspectives.  
Our work makes
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should benefit both Ceph and the HPC community
at large. 
%Throughout the evaluation, we observe that Ceph is an evolving
%technology under fast-paced development and shows great promises.

\end{abstract}