### 2013-11-06 cover.tex
Add sponsorship info

\thispagestyle{empty}

\hfill ORNL/TM-2013/151

\vspace{3em}

\begin{center}
National Center for Computatational Sciences
\end{center}

\vspace{3em}

\begin{center}
\textbf{\Large Ceph Parallel File System Evaluation Report}\\
\rule{5in}{1pt}%
\end{center}

\vspace{0.5in}

\begin{center}
\begin{tabular}{ c c }
Feiyi Wang  &  Mark Nelson \\ 
ORNL/NCCS & Inktank Inc. \\
\end{tabular}
\end{center}

\vspace{0.25in}

\begin{center}
Other contributors: \\
\vspace{1em}
Sarp Oral\\
Doug Fuller\\
Scott Atchley\\
Blake Caldwell\\
James Simmons\\
Brad Settlemyer\\
Jason Hill\\
Sage Weil (Inktank Inc.)
\end{center}

\vfill

\begin{center}
Prepared by \\
OAK RIDGE NATIONAL LABORATORY\\
Oak Ridge, Tennessee 37831-6283 \\ 
managed by \\
UT-BATTELLE, LLC \\
for the \\
U.S. DEPARTMENT OF ENERGY\\ 
\end{center}

This research was supported by, and used the resources of, the Oak Ridge
Leadership Computing Facility, located in the National Center for
Computational Sciences at ORNL, which is managed by UT Battelle, LLC for the
U.S. DOE (under the contract No. DE-AC05-00OR22725).

\pagebreak